---
title: "Introduction for R package SA23204161"
author: "Bosen Hu"
date: "2023-12-10"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Introduction for R package SA23204161}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
packages:
  - MASS
  - Rcpp
---

## Function G.nn.garrotte()

### Description

Solve the group linear regression using group non-negative garrotte method, given the linear regression data and the group information.

### Usage

G.nn.garrotte(x, y, g.index, p,normalized = TRUE, gram = TRUE)

### Arguments

|parameter|description|
|-----|-----|
|x|The design matrix, expressed as a matrix, has *n* rows and *m* columns.|
|y|The response vector, expressed as a matrix, has *n* rows and *1* column.|
|g.index|The group information, expressed as a vector, has *m* dimensions, start from 1, the same group has the same elements.|
|p|The group information, expressed as a vector, has *unique(g.index)* dimensions, the *i*-th element records the number of variables in the *i*-th group.|
|normalized|A logical variable, default is *TRUE*, scale every column of the design matrix.|
|gram|A logical variable, default is *TRUE*, apply gram-Schmidt orthogonal normalization for column vectors of the same group.|

### Value

A list containing the origin data *x*, *y*, *g.index*, *p*, and the shrinking matrix *d*, estimated parameters *betas*, solution path *path*.

### Example

```{r eval = FALSE}
library(MASS)
x<-mvrnorm(200,mu = rep(0,8),Sigma = diag(rep(1,8)))
eps<-rnorm(200,0,1)
beta<-matrix(c(.1,.2,.3,.3,.5,.9,.4,.4))
y<-x%*%beta+eps
g.index<-c(1,1,2,2,2,2,3,3)
p<-c(2,4,2)
fit<-G.nn.garrotte(x,y,g.index,p)
fit
```


## Function SNSMS()

### Description

Calculate the change-point that may occur in the time series data using the self-normalized sequential change-point detection method, which avoids calculating the long-run covariance matrix.

### Usage

SNSMS(L,x,theta,m,alpha,Tm,d)

### Arguments

|parameter|description|
|-----|-----|
|L|The object function, the dimension is same as *x*.|
|x|The time series data, can be a vector or a *d*-row matrix. Suppose the first *m* data are observed with no change points.|
|theta|Estimated parameter satisfying the sum of the first *m* object functions equal zero.|
|m|The number of observed data.|
|alpha|The significance level, taking value at 0.05 and 0.1.|
|Tm|the ratio of the monitoring horizon to the size of the training sample, taking value at 1, 2, 10 and >100.|
|d|The dimension of data, taking value at 1, 2 and 3.|

### Value

A numeric *k* representing the change point occurring at the time *m+k*.

### Example

```{r eval = FALSE}
L<-function(x,mu){(x-mu)^2}
x<-c(rnorm(200,0,1),rnorm(1800,2,1))
m<-180
theta<-mean(x[1:m])
alpha<-0.05
Tm<-10
d<-1
k<-SNSMS(L,x,theta,m,alpha,Tm,d)
plot(x,type="l")
abline(v=m+k,col="red")
```

```{r eval = FALSE}
library(MASS)
L<-function(x,mu){(x-mu)^2}
x<-cbind(t(mvrnorm(200,c(0,0),diag(c(1,1)))),t(mvrnorm(1800,c(1,2),diag(c(1,1)))))
m<-180
theta<-rowMeans(x[,1:m])
alpha<-0.05
Tm<-10
d<-2
SNSMS(L,x,theta,m,alpha,Tm,d)
```



## Function SNSMS_Cpp()

### Description

Calculate the change-point that may occur in the time series data using the self-normalized sequential change-point detection method, the Cpp version.

### Usage

SNSMS_Cpp(Function L,NumericMatrix x,NumericVector theta,
              int m,double alpha,int Tm,int d)

### Arguments

|parameter|description|
|-----|-----|
|L|The object function, the dimension is same as *x*.|
|x|The time series data, a *d*-row matrix. Suppose the first *m* data are observed with no change points.|
|theta|Estimated parameter satisfying the sum of the first *m* object functions equal zero.|
|m|The number of observed data.|
|alpha|The significance level, taking value at 0.05 and 0.1.|
|Tm|the ratio of the monitoring horizon to the size of the training sample, taking value at 1, 2, 10 and >100.|
|d|The dimension of data, taking value at 1, 2 and 3.|

### Value

A numeric *k* representing the change point occurring at the time *m+k*.

### Example

```{r eval = FALSE}
L<-function(x,mu){(x-mu)^2}
x<-matrix(c(rnorm(200,0,0.1),rnorm(1800,2,0.1)),nrow = 1)
m<-180
theta<-mean(x[1:m])
alpha<-0.05
Tm<-10
d<-1
k<-SNSMS_Cpp(L,x,theta,m,alpha,Tm,d)
plot(x[1,],type="l")
abline(v=m+k,col="red")
```

```{r eval = FALSE}
library(MASS)
L<-function(x,mu){(x-mu)^2}
x<-cbind(t(mvrnorm(200,c(0,0),diag(c(1,1)))),t(mvrnorm(1800,c(1,2),diag(c(1,1)))))
m<-180
theta<-rowMeans(x[,1:m])
alpha<-0.05
Tm<-10
d<-2
SNSMS_Cpp(L,x,theta,m,alpha,Tm,d)
```

## Dataset db.c

### Description

A (2,4,3)-dimension array. It gives the decision boundary when the significance level alpha, the monitoring scheme Tm and the dimension of samples are given.

### Usage

data(db.c)

### Format

The 1-th dimension: the significance level *alpha*.

The 2-th dimension: the monitoring scheme *Tm*.

The 3-th dimension: the dimension *d* of samples.

## References

Yuan M, Lin Y. Model selection and estimation in regression with grouped variables[J]. Journal of the Royal Statistical Society Series B: Statistical Methodology, 2006, 68(1): 49-67.

Chan N H, Ng W L, Yau C Y. A self-normalized approach to sequential change-point detection for time series[J]. Statistica Sinica, 2021, 31(1): 491-517.